---
title: "02b-Outlier-tests"
author: "Laura H Spencer"
date: "1/29/2022"
output: html_document
---

### Load libraries and source scripts 

```{r, message=FALSE, warning=FALSE, results=FALSE}
source("../references/biostats.R")

list.of.packages <- c("DESeq2", "RCurl", "tidyverse", "vegan", "pheatmap", "pastecs", "factoextra", "FactoMineR", "RColorBrewer", "tibble", "reshape2", "plotly", "corrplot", "PerformanceAnalytics", "cowplot", "here", "janitor", "clipr", "ggpubr", "forcats") #add new libraries here 
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# Load all libraries 
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X)) 
})

# Need these packages for SNP assessment 
# if (!requireNamespace("BiocManager", quietly=TRUE))
#     install.packages("BiocManager")
# BiocManager::install("gdsfmt")
# BiocManager::install("SNPRelate")
```


### Import sample info, library/file names, and then join 

```{r}
library.stats <- read.csv("../data/library-prep-stats.csv", header=T) %>% clean_names %>%
  mutate_at(vars(tank, p_h, batch, date_extracted), as.factor)

sample.info <- read.csv("../data/RCK-exp-design.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(Tank, Treatment, Treatment_Tank), as.factor) %>%
  left_join( ., library.stats, by=c("Sample"="tank_crab"))
```

### Import counts matrix file as dataframe, extract sample # from file/column names to simplify them 

```{r}
# FOR v6 ONLY: (DIFFERENT FILE NAMES)
# counts <- data.frame(read.table("../results/featureCounts/featurecounts_redking_gene-v6", header = T, stringsAsFactors = F, fill = FALSE))
# counts <- counts %>%  column_to_rownames(var="Geneid")
# counts <- counts %>% rename_all(~as.character(str_sub(colnames(counts)) %>% 
#                              gsub("X.scratch.lspencer.2022.redking.OA.aligned.bowtie2.sorted.", "", .) %>% 
#                              gsub(".sorted.bam", "", .)))
```

### Summarize counts and visualize (remove last column - that's undetermined counts)

```{r}
LOO.res <- setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("counts.total", "genes.total", "genes.ave", "genes.dropped", "genes.remain", "diff.AmbMod", "diff.AmbLow", "Diff.ModLow"))
LOO.PCA <- list()

for (i in 1:nrow(sample.info)) {

sample.info <- read.csv("../data/RCK-exp-design.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(Tank, Treatment, Treatment_Tank), as.factor)

counts <- data.frame(read.table("../results/featureCounts/featurecounts_redking_gene-v1", header = T, stringsAsFactors = F, fill = FALSE))
counts <- counts %>%  column_to_rownames(var="Geneid")
counts <- counts %>% rename_all(~as.character(str_sub(colnames(counts)) %>%
                             gsub("X.scratch.lspencer.2022.redking.OA.aligned.bowtie.sorted.deduped.", "", .) %>%
                             gsub(".sorted.bam", "", .)))

## FILTER WHOLE SAMPLES. 
remove.list <- sample.info$Sample[i]
counts <- counts[ , -which(names(counts) %in% remove.list)]
sample.info <- sample.info[ -which(sample.info$Sample %in% remove.list), ]

### Remove extraneous gene info, then transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
counts.t <- t(counts[,-1:-5]) #remove extraneous columns, transform data to have each sample a row, each column a gene 

keep <- colSums(counts.t) >= 10
counts.ts <- counts.t[,keep]

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk <- merge(x=sample.info[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

# Run DESeq2 analysis
dds.pH <- DESeqDataSetFromMatrix(countData = counts.tk[,grepl("evm", colnames(counts.tk))] %>% t(),
                              colData = counts.tk[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)
vsd.pH <- varianceStabilizingTransformation(dds.pH, blind=FALSE)
dds.DESeq.pH <- DESeq(dds.pH) 

res.pco2.AM <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Moderate"), alpha=0.05)
res.pco2.AL <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Low"), alpha=0.05)
res.pco2.ML <- results(dds.DESeq.pH, contrast=c("Treatment", "Moderate", "Low"), alpha=0.05)


LOO.res[i,1] <- sum(colSums(counts %>% select(contains("Tank"))))
LOO.res[i,2] <- nrow(counts)
LOO.res[i,3] <- mean(colSums(counts %>% select(contains("Tank")) != 0))
LOO.res[i,4] <- ncol(counts.t) - ncol(counts.ts)
LOO.res[i,5] <- ncol(counts.ts)
LOO.res[i,6] <- sum(res.pco2.AM$padj < 0.05, na.rm=TRUE)
LOO.res[i,7] <- sum(res.pco2.AL$padj < 0.05, na.rm=TRUE)
LOO.res[i,8] <- sum(res.pco2.ML$padj < 0.05, na.rm=TRUE)

p <- ggplotly(
  plotPCA(vsd.pH, intgroup="Treatment") + 
           ggtitle("PCA by Treatment (var-stabilizing transformed)") + 
    geom_point(size=3, aes(text=colnames(vsd.pH))) + 
    theme_minimal()+ stat_ellipse(), tooltip = "text")
LOO.PCA[[i]] <- p
}
LOO.PCA

LOO.res[,9] <- sample.info$Sample
LOO.res <- LOO.res %>% left_join(sample.info[c("Sample", "Treatment")], by=c("V9"="Sample"))

LOO.stats <- list()
for (i in 1:8) {
  p <-  ggplotly(ggplot(data = LOO.res) +
           geom_bar(aes_string(x=noquote(names(LOO.res)[9]), 
                               y=noquote(names(LOO.res)[i]),
                               fill=noquote(names(LOO.res)[10]),
                               text=noquote(names(LOO.res)[9])), stat = "identity") + 
    ggtitle(colnames(LOO.res)[i]) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)), 
    tooltip = "text")
LOO.stats[[i]] <- p  
}
LOO.stats

#LOO.res[c(colnames(LOO.res[i]))]

LOO.res %>% write_clip()
```



```{r}
sample.info$Sample
```
## Inspect DEGs dropped when Tank 7 Crab 4 is removed. 

```{r}
sample.info <- read.csv("../data/RCK-exp-design.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(Tank, Treatment, Treatment_Tank), as.factor)
counts <- data.frame(read.table("../results/featureCounts/featurecounts_redking_gene-v1", header = T, stringsAsFactors = F, fill = FALSE))
counts <- counts %>%  column_to_rownames(var="Geneid")
counts <- counts %>% rename_all(~as.character(str_sub(colnames(counts)) %>%
                             gsub("X.scratch.lspencer.2022.redking.OA.aligned.bowtie.sorted.deduped.", "", .) %>%
                             gsub(".sorted.bam", "", .)))


# Generate DEG list with no samples removed 

### Remove extraneous gene info, then transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
counts.t <- t(counts[,-1:-5]) #remove extraneous columns, transform data to have each sample a row, each column a gene 

keep <- colSums(counts.t) >= 10
counts.ts <- counts.t[,keep]

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk <- merge(x=sample.info[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

# Run DESeq2 analysis
dds.pH <- DESeqDataSetFromMatrix(countData = counts.tk[,grepl("evm", colnames(counts.tk))] %>% t(),
                              colData = counts.tk[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)

#vsd.pH <- varianceStabilizingTransformation(dds.pH, blind=FALSE)
dds.DESeq.pH <- DESeq(dds.pH) 

res.pco2.AM <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Moderate"), alpha=0.05)
res.pco2.AL <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Low"), alpha=0.05)
res.pco2.ML <- results(dds.DESeq.pH, contrast=c("Treatment", "Moderate", "Low"), alpha=0.05)

# =======================================
# Not re-do that with Tank7.Crab4 removed 

## FILTER WHOLE SAMPLES. 
remove.list <- "Tank_7_Crab_4"
counts.noout <- counts[ , -which(names(counts) %in% remove.list)]
sample.info.noout <- sample.info[ -which(sample.info$Sample %in% remove.list), ]

# Remove extraneous gene info, then transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
counts.t.noout <- t(counts[,-1:-5]) #remove extraneous columns, transform data to have each sample a row, each column a gene 

keep.noout <- colSums(counts.t.noout) >= 10
counts.ts.noout <- counts.t.noout[,keep.noout]

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk.noout <- merge(x=sample.info.noout[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts.noout, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

# Run DESeq2 analysis
dds.pH.noout <- DESeqDataSetFromMatrix(countData = counts.tk.noout[,grepl("evm", colnames(counts.tk.noout))] %>% t(),
                              colData = counts.tk.noout[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)

#vsd.pH <- varianceStabilizingTransformation(dds.pH, blind=FALSE)
dds.DESeq.pH.noout <- DESeq(dds.pH.noout) 

res.pco2.AM.noout <- results(dds.DESeq.pH.noout, contrast=c("Treatment", "Ambient", "Moderate"), alpha=0.05)
res.pco2.AL.noout <- results(dds.DESeq.pH.noout, contrast=c("Treatment", "Ambient", "Low"), alpha=0.05)
res.pco2.ML.noout <- results(dds.DESeq.pH.noout, contrast=c("Treatment", "Moderate", "Low"), alpha=0.05)

# ===============
# Not identify genes that aren't on both results lists 

setdiff(rownames(subset(res.pco2.AM, padj < 0.05)),
        rownames(subset(res.pco2.AM.noout, padj < 0.05))) %>% length()

setdiff(rownames(subset(res.pco2.AL, padj < 0.05)),
        rownames(subset(res.pco2.AL.noout, padj < 0.05))) %>% length()

setdiff(rownames(subset(res.pco2.ML, padj < 0.05)),
        rownames(subset(res.pco2.ML.noout, padj < 0.05))) %>% length()


test <- setdiff(rownames(subset(res.pco2.AL, padj < 0.05)),
        rownames(subset(res.pco2.AM.noout, padj < 0.05)))

diff_plots_AL = list()
for (i in 1:length(test)) {
a <-  plotCounts(dds.DESeq.pH, gene=test[i], intgroup=c("Treatment"), returnData = TRUE)
b <- ggplot(a %>% filter(Treatment==c("Ambient", "Low")) 
                         %>% rownames_to_column("sample"),
       aes(x=Treatment, y=count, color=Treatment, label=sample)) +
  geom_point(position=position_jitter(w = 0.1,h = 0)) +
    geom_text() +
    theme_bw() +
    ggtitle(test[i]) +
    theme(plot.title = element_text(hjust = 0.5))
diff_plots_AL[[i]] <- b
}
diff_plots_AL[1:10] #10 of many 
```

# Inspect outlier genes and test removing those genes 

Interesting that there are no outlier genes identified in any of the above Diff. Expression results. But what about whole samples? DESeq2 calculates Cook's distances to aid in outlier detection. 

```{r}
sample.info <- read.csv("../data/RCK-exp-design.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(Tank, Treatment, Treatment_Tank), as.factor)
counts <- data.frame(read.table("../results/featureCounts/featurecounts_redking_gene-v1", header = T, stringsAsFactors = F, fill = FALSE))
counts <- counts %>%  column_to_rownames(var="Geneid")
counts <- counts %>% rename_all(~as.character(str_sub(colnames(counts)) %>%
                             gsub("X.scratch.lspencer.2022.redking.OA.aligned.bowtie.sorted.deduped.", "", .) %>%
                             gsub(".sorted.bam", "", .)))


# Generate DEG list with no samples removed 

### Remove extraneous gene info, then transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
counts.t <- t(counts[,-1:-5]) #remove extraneous columns, transform data to have each sample a row, each column a gene 

keep <- colSums(counts.t) >= 10
counts.ts <- counts.t[,keep]

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk <- merge(x=sample.info[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

# Run DESeq2 analysis
dds.pH <- DESeqDataSetFromMatrix(countData = counts.tk[,grepl("evm", colnames(counts.tk))] %>% t(),
                              colData = counts.tk[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)
dds.DESeq.pH <- DESeq(dds.pH) 

res.pco2.AM <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Moderate"), alpha=0.05)
res.pco2.AL <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Low"), alpha=0.05)
res.pco2.ML <- results(dds.DESeq.pH, contrast=c("Treatment", "Moderate", "Low"), alpha=0.05)

# Generate summary stats 
sum(colSums(counts %>% select(contains("Tank"))))
nrow(counts)
mean(colSums(counts %>% select(contains("Tank")) != 0))
ncol(counts.t) - ncol(counts.ts)
ncol(counts.ts)
sum(res.pco2.AM$padj < 0.05, na.rm=TRUE)
sum(res.pco2.AL$padj < 0.05, na.rm=TRUE)
sum(res.pco2.ML$padj < 0.05, na.rm=TRUE)
```
## Now identify genes that have outlier counts in T7C4 

```{r}
#Here is a boxplot of the Cook’s distances. Sample T7C4 slightly higher Cooks distances, but not much. 
par(mar=c(8,5,2,2))
boxplot(log10(assays(dds.DESeq.pH)[["cooks"]]), range=0, las=2, cex.axis=0.6)

#ggplotly(
  ggplot(dat) + 
  geom_boxplot(aes(x = ind, y = values)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))#)

# Inspect outliers according to the boxplot of T7C4 Cook's distances:   
(assays(dds.DESeq.pH)[["cooks"]] %>% as.data.frame())$Tank_7_Crab_4 %>% log2() %>% boxplot()

# Extract outliers using car package
cooks <- (assays(dds.DESeq.pH)[["cooks"]] %>% as.data.frame())
outliers<-Boxplot(~Tank_7_Crab_4, data=cooks, las=2, cex.axis=0.6, cex=0.5, id=list(n=Inf)) 

#How many outliers?
length(outliers)

## can also use this code if I want to identify outliers by factor or by sample 
# cooks <- (assays(dds.DESeq.pH)[["cooks"]] %>% as.data.frame()) %>% rownames_to_column(var="gene") %>% 
#   pivot_longer(cols = starts_with("Tank"), names_to = "Sample", values_to = "Cooks") %>% 
#   left_join(sample.info[,c("Sample", "Treatment")]) %>% 
#   mutate(Sample=as.factor(Sample))
# outliers<-Boxplot(Cooks~Sample, data=cooks, las=2, cex.axis=0.6, cex=0.5, id=list(n=Inf))

diff_plots = list()
for (i in 1:length(outliers[1:32])) {
a <-  plotCounts(dds.DESeq.pH, gene=outliers[i], intgroup=c("Treatment"), returnData = TRUE)
b <- ggplot(a %>% rownames_to_column("sample"),
       aes(x=Treatment, y=count, color=Treatment, label=sample)) +
  geom_text(size=2) +
    theme_bw(base_size = 6) +
    ggtitle(test[i]) +
    theme(plot.title = element_text(hjust = 0.5))
diff_plots[[i]] <- b
}
diff_plots

# Save some figures to pdf 
fig <- ggarrange(plotlist=diff_plots, 
                       ncol=4, nrow=4,
                       common.legend = TRUE, legend="top")

pdf("../results/deseq2/some-outlier-genes.pdf", onefile = TRUE)
fig
dev.off()
```
## Re-run analysis with outlier GENES removed 

```{r}
sample.info <- read.csv("../data/RCK-exp-design.csv", header=T, na.strings = NA) %>%
          mutate_at(vars(Tank, Treatment, Treatment_Tank), as.factor)
counts <- data.frame(read.table("../results/featureCounts/featurecounts_redking_gene-v1", header = T, stringsAsFactors = F, fill = FALSE))
counts <- counts %>%  column_to_rownames(var="Geneid")
counts <- counts %>% rename_all(~as.character(str_sub(colnames(counts)) %>%
                             gsub("X.scratch.lspencer.2022.redking.OA.aligned.bowtie.sorted.deduped.", "", .) %>%
                             gsub(".sorted.bam", "", .)))

# REMOVE OUTLIER GENES 
counts <- counts[-which(rownames(counts) %in% outliers),]

### Remove extraneous gene info, then transpose dataframe so each row = a sample (aka "objects"), and each column = genes (aka "variables") 
counts.t <- t(counts[,-1:-5]) #remove extraneous columns, transform data to have each sample a row, each column a gene 

keep <- colSums(counts.t) >= 10
counts.ts <- counts.t[,keep]

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk <- merge(x=sample.info[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

# Run DESeq2 analysis
dds.pH <- DESeqDataSetFromMatrix(countData = counts.tk[,grepl("evm", colnames(counts.tk))] %>% t(),
                              colData = counts.tk[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)
dds.DESeq.pH <- DESeq(dds.pH) 

res.pco2.AM <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Moderate"), alpha=0.05)
res.pco2.AL <- results(dds.DESeq.pH, contrast=c("Treatment", "Ambient", "Low"), alpha=0.05)
res.pco2.ML <- results(dds.DESeq.pH, contrast=c("Treatment", "Moderate", "Low"), alpha=0.05)

# Generate summary stats 
sum(colSums(counts %>% select(contains("Tank"))))
nrow(counts)
mean(colSums(counts %>% select(contains("Tank")) != 0))
ncol(counts.t) - ncol(counts.ts)
ncol(counts.ts)
sum(res.pco2.AM$padj < 0.05, na.rm=TRUE)
sum(res.pco2.AL$padj < 0.05, na.rm=TRUE)
sum(res.pco2.ML$padj < 0.05, na.rm=TRUE)

# How does the PCA look with those outlier genes removed? Very similar. 
vsd.pH <- varianceStabilizingTransformation(dds.pH, blind=FALSE)
ggplotly(
  plotPCA(vsd.pH, intgroup="Treatment") + 
           ggtitle("PCA by Treatment (var-stabilizing transformed)") + 
    geom_point(size=3, aes(text=colnames(vsd.pH))) + 
    theme_minimal()+ stat_ellipse(), tooltip = "text")
```

# DECISION: REMOVE OUTLIER SAMPLE 
This is the most straight-forward, defensible option, AND it really does appear to be an outlier that influences the DEG results. 


## How many consensus DEGs are there among pH treatments using Bowtie2 and STAR aligners? 

```{r}
# Total number of genes in both dataset for analysis: 
intersect(rownames(res.pco2.Bowtie), rownames(res.pco2.Star)) %>% length()

# Ambient vs. Moderate:  Bowtie2=687 DEGs, STAR=630 DEGs, consensus DEGs=376 DEGs 
res.pco2.AM.Bowtie %>% nrow()
res.pco2.AM.Star %>% nrow()
intersect(rownames(res.pco2.AM.Bowtie), rownames(res.pco2.AM.Star)) %>% length()

# Ambient vs. Low:  Bowtie2=1491 DEGs, STAR=966 DEGs, consensus DEGs=633 DEGs 
res.pco2.AL.Bowtie %>% nrow()
res.pco2.AL.Star %>% nrow()
intersect(rownames(res.pco2.AL.Bowtie), rownames(res.pco2.AL.Star)) %>% length()

# Moderate vs. Low:  Bowtie2=15 DEGs, STAR=19 DEGs, consensus DEGs=9 DEGs 
res.pco2.ML.Bowtie %>% nrow()
res.pco2.ML.Star %>% nrow()
intersect(rownames(res.pco2.ML.Bowtie), rownames(res.pco2.ML.Star)) %>% length()
```


Enrichment analysis
```{r}
# DEGs between Ambient & Moderate 
P.plat.blast %>% filter(geneID %in% c(intersect(rownames(res.pco2.AM.Bowtie), rownames(res.pco2.AM.Star)))) %>% as_tibble() %>%
  dplyr::select("SPID") %>% na.omit() %>% unlist() %>% as.vector() %>% write_clip()

# DEGs between Ambient & Low
P.plat.blast %>% filter(geneID %in% c(intersect(rownames(res.pco2.AL.Bowtie), rownames(res.pco2.AL.Star)))) %>% as_tibble() %>%
  dplyr::select("SPID") %>% na.omit() %>% unlist() %>% as.vector() %>% write_clip()

# DEGs between Moderate & Low (includes just 1 annotated genes) 
P.plat.blast %>% filter(geneID %in% c(intersect(rownames(res.pco2.ML.Bowtie), rownames(res.pco2.ML.Star)))) %>% as_tibble() %>%
  dplyr::select("SPID") %>% na.omit() %>% unlist() %>% as.vector() %>% write_clip()

# All genes (background, includes 6,508 annotated genes identified by both STAR and Bowtie2)
P.plat.blast %>% filter(geneID %in% c(intersect(rownames(res.pco2.Bowtie), rownames(res.pco2.Star)))) %>% as_tibble() %>%
  dplyr::select("SPID") %>% na.omit() %>% unlist() %>% as.vector() %>% write_clip()

# Read in DAVID results
# Between ambient and moderate pH 
read_delim(file="../results/DAVID-Enriched-BP_Amb-Mod_no-outlier-T7C4_consensus.txt", delim = "\t") %>% 
  mutate(GO = str_extract(Term, "GO(.*?)~")) %>% 
  mutate(GO = gsub("~", "", GO)) %>% dplyr::select(GO, PValue) %>% na.omit() %>% write_clip()

# Between ambient and low pH 
read_delim(file="../results/DAVID-Enriched-BP_Amb-Low_no-outlier-T7C4_consensus.txt", delim = "\t") %>% 
  mutate(GO = str_extract(Term, "GO(.*?)~")) %>% 
  mutate(GO = gsub("~", "", GO)) %>% dplyr::select(Term) %>% na.omit() %>% write_clip()
```

