---
title: "xx-WGCNA"
author: "Laura H Spencer"
date: "2/28/2022"
output: html_document
---


```{r, message=FALSE, warning=FALSE, results=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- c("tidyverse", "RColorBrewer", "ggpmisc")
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- c("WGCNA", "DESeq2")

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})
```

```{r}
# load data
load(file = "../results/gene-counts") # object is counts
load(file = "../results/gene-counts-trans") #object is counts.t 
load(file="../data/sample.info") #object = sample.info
load(file = "../results/bowtie/per-gene-summary-stats")

keep1 <- colMeans(counts.t, na.rm=TRUE) >= 10 #identify genes with mean count across all samples (excluding NAs = 10)
keep2 <- rowSums( counts >= 10 ) >= 0.2*43 #identify genes with counts>=10 across at minimum 25% of the samples
keep <- Reduce(intersect, list(names(which(keep1 == T)), names(which(keep2 == T)))) # list of genes meeting both of the above criteria 
counts.ts <- counts.t[,keep]

print(paste("# genes remaining after pre-filtering:", ncol(counts.ts)))
print(paste("# of genes dropped:", ncol(counts.t) - ncol(counts.ts), sep=" "))

# merge count data with sample key, reset row names as sample names, and arrange by infection, then temperature, then day 
counts.tk <- merge(x=sample.info[,c("Sample", "Tank", "Treatment", "Treatment_Tank")], by.x="Sample", y=counts.ts, by.y="row.names") %>% 
  arrange(Treatment, Tank)  %>% column_to_rownames(var="Sample") %>% droplevels()

all(rownames(counts.tk) == counts.tk %>% dplyr::select(starts_with("evm")) %>% t() %>% colnames()) #check that rownames of untransformed matrix match column names of transformed matrix. Should print 'TRUE' 

dds.pH <- DESeqDataSetFromMatrix(countData = counts.tk[,grepl("evm", colnames(counts.tk))] %>% t(),
                              colData = counts.tk[,"Treatment", drop=FALSE] ,
                              design = ~ Treatment)

vsd.pH <- varianceStabilizingTransformation(dds.pH, blind=FALSE)

counts.trans <- assay(vsd.pH) %>% t() %>% as.matrix() %>% as.data.frame() %>%
  rownames_to_column(var = "Sample") %>% 
  left_join(sample.info[,c("Sample", "Treatment", "Tank", "Treatment_Tank")]) %>%
  column_to_rownames(var = "Sample") %>% 
  select(Treatment, Tank, Treatment_Tank, contains("evm")) %>% 
  mutate(Treatment=factor(Treatment, levels=c("Ambient", "Moderate", "Low")))
save(counts.trans, file = "../results/wgcna/counts.filt")
```

Format of expression data should be: 
- rownames = sample 
- columns = genes 
- normalized 

```{r}
# Use WGCNA's `goodSamplesGenes` to remove any genes with very low variability or missing data  
datExpr0 <- counts.trans[,-1:-3]
gsg = goodSamplesGenes(datExpr0, verbose = 3);
sum(gsg$goodGenes==FALSE)  #were any removed?  Nope. 
ncol(gsg$allOK)

if (!gsg$allOK)
{
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0) 
     printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0) 
     printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}
```

```{r}
sampleTree = hclust(dist(datExpr0), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
# The user should change the dimensions if the window is too large or too small.
sizeGrWindow(12,9)
#pdf(file = "Plots/sampleClustering.pdf", width = 12, height = 9);
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)


# Plot a line to show the cut
abline(h = 100, col = "red");
# Determine cluster under the line
clust = cutreeStatic(sampleTree, cutHeight = 100, minSize = 10)
table(clust)
# clust 1 contains the samples we want to keep.
keepSamples = (clust==1)
datExpr = datExpr0[keepSamples, ]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)

```

```{r}
# traitData = read.csv("wgcna-testing/ClinicalTraits.csv");
# dim(traitData)
# names(traitData)


# remove columns that hold information we do not need.
allTraits = counts.trans[, 1:3] %>% rownames_to_column("Sample") %>%
  mutate(pH=Treatment) %>%
  mutate_at("pH", function(x) {
    case_when(
      x == "Ambient" ~ 8.0,
      x == "Moderate" ~ 7.8,
      x == "Low" ~ 7.6)})

dim(allTraits)
names(allTraits)

# # Form a data frame analogous to expression data that will hold the traits.
# 
Samples = rownames(datExpr);
traitRows = match(Samples, allTraits$Sample);
datTraits = allTraits[traitRows, -1];
rownames(datTraits) = allTraits[traitRows, 1];

# 
collectGarbage();
```

```{r}
# Re-cluster samples
sampleTree2 = hclust(dist(datExpr), method = "average")
# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = numbers2colors(sapply(datTraits,as.numeric), signed = FALSE, colors = colorRampPalette(brewer.pal(8, "Set2"))(14));

# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTraits), 
                    main = "Sample dendrogram with treatment and tank ids")
```


```{r}
#save(datExpr, datTraits, file = "wgcna-testing/FemaleLiver-01-dataInput.RData")
```

```{r}
options(stringsAsFactors = FALSE);
#lnames = load(file = "../wgcna-testing/FemaleLiver-01-dataInput.RData");
```

"Constructing a weighted gene network entails the choice of the soft thresholding power β to which co-expression
similarity is raised to calculate adjacency. The authors of [1] have proposed to choose the soft thresholding power based on the criterion of approximate scale-free topology." 

I will choose the power 9, which is the lowest power for which the scale-free topology fit
index reaches 0.90.

```{r}
# Choose a set of soft-thresholding powers
powers = c(c(1:14), seq(from = 16, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5, networkType = "signed")
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```
I now calculate the adjacencies, using the soft thresholding power 9:

```{r}
softPower = 14;
adjacency = adjacency(datExpr, power = softPower, type="signed", corFnc="bicor");

# for testing purposes, subset expression dataframe 
#adjacency = adjacency(datExpr[,1:2000], power = softPower, type="signed", corFnc="bicor");
```

To minimize effects of noise and spurious associations, we transform the adjacency into Topological Overlap Matrix, and calculate the corresponding dissimilarity:

```{r}
TOM = TOMsimilarity(adjacency);
dissTOM = 1-TOM
save(dissTOM, file = "../results/wgcna/dissTom")
```
We now use hierarchical clustering to produce a hierarchical clustering tree (dendrogram) of genes. Note that we use the function hclust that provides a much faster hierarchical clustering routine than the standard hclust function.

```{r}
# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average");
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04);
```

```{r}
# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 50;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                deepSplit = 2, pamRespectsDendro = FALSE,
                minClusterSize = minModuleSize);
table(dynamicMods)
```
```{r}
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene clustering on TOM-based dissimilarity\nDendrogram and module colors")
```

The Dynamic Tree Cut may identify modules whose expression profiles are very similar. It may be prudent to merge
such modules since their genes are highly co-expressed. To quantify co-expression similarity of entire modules, we
calculate their eigengenes and cluster them on their correlation.

We choose a height cut of 0.25, corresponding to correlation of 0.75, to merge (see Fig. 4):


```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

MEDissThres = 0.25
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")
# Call an automatic merging function
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;

```


```{r}
#sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene clustering on TOM-based dissimilarity\nDendrogram and module colors")
#dev.off()
#sizeGrWindow(12, 9)
plotDendroAndColors(geneTree, mergedColors,
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene clustering on TOM-based dissimilarity\nFinalized modules")
```



```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs;
# Save module colors and labels for use in subsequent parts
#save(MEs, moduleLabels, moduleColors, geneTree, file = "../results/wgcna/networkConstruction-stepByStep.RData")
```


```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits[,c("pH"), drop=F], use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

```

```{r}
sizeGrWindow(10,6)
# Will display correlations and their p-values
# textMatrix =  paste(signif(moduleTraitCor, 2), "\n(",
#                            signif(moduleTraitPvalue, 1), ")", sep = "");
textMatrix =  paste(signif(moduleTraitCor, 2), " (",
                           signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(datTraits[,c("pH"), drop=F]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = greenWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.75,
               cex.main = 1.0,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```


```{r}
# Define variable pH containing the weight column of datTrait
pH = as.data.frame(datTraits$pH);
names(pH) = "pH"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneTraitSignificance = as.data.frame(cor(datExpr, pH, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(pH), sep="");
names(GSPvalue) = paste("p.GS.", names(pH), sep="");
```


```{r}
annot = P.plat.bestblast %>% filter(id %in% names(datExpr))
dim(annot)
names(annot)
probes = names(datExpr)
probes2annot = match(probes, annot$id)
# The following is the number or probes without annotation:
sum(is.na(probes2annot))
# Should return 0.
```

```{r}
# Create the starting data frame
geneInfo0 = data.frame(id = probes,
                      hit_acc = annot$hit_acc[probes2annot],
                      hit_desc = annot$hit_desc[probes2annot],
                      moduleColor = moduleColors,
                      geneTraitSignificance,
                      GSPvalue)
# Order modules by their significance for pH
modOrder = order(-abs(cor(MEs, pH, use = "p")));
# Add module membership information in the chosen order
for (mod in 1:ncol(geneModuleMembership))
{
  oldNames = names(geneInfo0)
  geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]], 
                         MMPvalue[, modOrder[mod]]);
  names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
                       paste("p.MM.", modNames[modOrder[mod]], sep=""))
}
# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$moduleColor, -abs(geneInfo0$GS.pH));
geneInfo = geneInfo0[geneOrder, ]
#write.csv(geneInfo, file = "../results/wgcna/geneInfo.csv")
```


# departure from WGCNA tutorials 

## Generate line plots of genes for each module of interest

```{r}
# Generate line plots of significant modules showing only genes significantly correlated to module 
# List of modules (their color IDs) that are significantly associated with pH 
modules <- moduleTraitPvalue %>% as.data.frame() %>% filter(pH<0.05) %>% rownames() %>% str_remove(pattern = "ME")
modules.p <- paste("p.MM.", modules, sep = "")
modules.cor <- noquote(paste("MM.", modules, sep = ""))

# create empty list to store summary statistics for each modules' sign. genes 
module.stats <- vector("list", length(modules))
names(module.stats) <- modules

# create empty list to store line plots for each modules' sign. genes 
module.lineplots <- vector("list", length(modules))
names(module.lineplots) <- modules

for (i in 1:length(modules)) {

a <- geneInfo %>%
  filter(moduleColor==modules[i]) %>% 
  filter(p.GS.pH<0.05) %>% filter(!!as.symbol(modules.p[i])<0.05) %>% 
  select(id, moduleColor, GS.pH, p.GS.pH, !!as.symbol(modules.p[i]), !!as.symbol(modules.cor[i])) %>% 
  left_join(genes.stats, by = c("id"="gene")) %>% 
  #group_by(Treatment, moduleColor) %>% mutate(mean_z = scale(mean)) %>% ungroup() %>% #use this line of code to z-transform 
  #filter(moduleColor %in% modules) %>% 
  mutate(pH=Treatment) %>%
  mutate_at("pH", function(x) {
    case_when(
      x == "Ambient" ~ 8.0,
      x == "Moderate" ~ 7.8,
      x == "Low" ~ 7.6)})

formula <- y ~ x

module.lineplots[[i]] <-  ggplot(a, aes(x=pH, y=median)) + geom_line(aes(group=id), size=.01, color=modules[i]) +
  theme_minimal() + facet_wrap(~moduleColor + change) + 
  geom_smooth(method=lm, formula = formula, se = FALSE, color="blue", size=1.5) +
  scale_x_continuous(breaks=c(7.6, 7.8, 8.0)) + 
  theme(plot.title = element_text(size=10), axis.text.x = element_text(hjust = 0.2)) +
  xlab("pH Treatment") +
  stat_poly_eq(aes(label = ..eq.label..), 
               label.x.npc = 0.5, label.y.npc = 0.99,
               formula = formula, parse = TRUE, size = 3)

module.stats[[i]] <- a %>% select(-Treatment, -mean, -sd, -cv, -pH) %>% distinct()
}
print(module.lineplots)

# Save some figures to pdf 
fig <- ggarrange(plotlist=module.lineplots, 
                       ncol=2, nrow=2,
                       common.legend = TRUE, legend="top")
pdf("../results/wgcna/pH-associated-modules-genes.pdf", onefile = TRUE)
fig
dev.off()
```


#### Get kME

kME calculates module membership, and is used in the subsequent GO-MWU analysis that we'll perform on significant modules. 

We'll keep things simple by making a table of the kME for all genes. In the actual GO-MWU analysis, we'll set the kME of all genes that aren't in the significant module to 0. But no need to worry about that now!

```{r membership_and_gene_info}
kME_table <- signedKME(CrabExpr0,
                       datME = MEs0)
# Write kME file
fileName <- paste0(file_start, "kME_table.txt")
write.table(kME_table, file = fileName,
            sep = "\t",
            row.names = TRUE,
            col.names = TRUE)
```

-------------------------

# other stuff 

```{r}
names(datExpr)[moduleColors=="brown"]
module = "salmon"
column = match(module, modNames);
moduleGenes = moduleColors==module;

sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for pH",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```



```{r}
# Here I try to make lineplots with delta counts (from ambient) - very ugly and not informative, but this code might be useful
genes.stats %>% ungroup() %>%
  filter(gene %in% names(datExpr)[moduleColors==module]) %>% 
  select(gene, Treatment, mean, hit_acc, hit_desc, SPID) %>%
  pivot_wider(names_from = "Treatment", values_from = "mean") %>% 
  mutate(amb=0, mod=Moderate-Ambient, low=Low-Ambient) %>%
  select(-Ambient, -Moderate, -Low) %>%
  pivot_longer(cols = c(amb, mod, low), names_to = "pH", values_to = "delta") %>%
  mutate_at("pH", function(x) {
    case_when(
      x == "amb" ~ 8.0,
      x == "mod" ~ 7.8,
      x == "low" ~ 7.6)}) %>% 
  ggplot(aes(x=pH, y=delta, group=gene)) + geom_line(size=.01) +
  theme_minimal() + #facet_wrap(~change) +
  scale_x_continuous(breaks=c(7.6, 7.8, 8.0)) +
  theme(plot.title = element_text(size=10), axis.text.x = element_text(hjust = 0.2)) +
  ylab("Per-treatment mean of normalized counts") + ggtitle("Abundances of DEGs by treatment & relative expression") 
```


```{r}

genes.stats %>% 
  left_join(geneInfo %>% select(id, moduleColor, GS.pH, p.GS.pH), by = c("gene"="id")) %>% 
  #group_by(Treatment, moduleColor) %>% mutate(mean_z = scale(mean)) %>% ungroup() %>% #use this line of code to z-transform 
  filter(moduleColor %in% modules) %>% 
  mutate(pH=Treatment) %>%
  mutate_at("pH", function(x) {
    case_when(
      x == "Ambient" ~ 8.0,
      x == "Moderate" ~ 7.8,
      x == "Low" ~ 7.6)}) %>% 
  ggplot(aes(x=pH, y=log10(mean))) + geom_line(aes(group=gene), size=.01) +
  theme_minimal() + facet_wrap(~moduleColor+change) + geom_smooth(method=lm, se = FALSE, color="red", size=1.5) +
  scale_x_continuous(breaks=c(7.6, 7.8, 8.0)) + 
  theme(plot.title = element_text(size=10), axis.text.x = element_text(hjust = 0.2)) +
  xlab("Per-treatment mean of normalized counts") + ggtitle("Abundances of DEGs by treatment & relative expression")

test <- genes.stats %>% filter(gene %in% names(datExpr)[moduleColors==module]) %>% 
  mutate(pH=Treatment) %>%
  mutate_at("pH", function(x) {
    case_when(
      x == "Ambient" ~ 8.0,
      x == "Moderate" ~ 7.8,
      x == "Low" ~ 7.6)})
hist(log(test$mean))
shapiro.test(log(test$mean))
summary(lm(log(mean)~pH, data=test))

# Try plotting expression level of all genes for all samples, organized by pH 
modules
```

